I am grateful to my Ph.D. advisor Bala Krishnamoorthy, Professor,  Department of mathematics and statistics for his selfless support and patience as an academic advisor. As my research mentor, I have learned from him the importance of critical thinking. I have the uttermost respect for him for giving me complete freedom to approach research problems. I am also very thankful to him for providing me financial support for past several months.  

Special thanks to Kevin Vixie for help during the summer for GQE preparation. During that time I learned a lot about writing proofs, even though I did my undergrad and master's in mechanical engineering. 

Special thanks to Gregory Dreficus for helping me to land an internship at MDF, ORNL. This was the turning point for me to pick $3$D printing for my Ph.D. thesis. He is also a great friend and collaborator. 

I would also like to thanks Brian Post and Lonnie Love for giving me the opportunity to work as an intern at MDF. 

I would also like to thanks Narasimha Boddeti, assistant professor at the School of Mechanical and Materials Engineering, and Yiran Guo for their invaluable support during our collaboration.
I would also like to thanks Enrique Alvarado for useful discussions about my work.

\newpage
\textbf{Most importantly}, I would like to thanks, my parents and sister. Doing a Ph.D. is a solo journey. You enter into uncharted territory and trying to figure your way out. They have helped me a lot in this journey, by their selfless support and motivation when I was feeling low.
